\chapter{Application Security}
\label{ch:application-security}

\section{Decentralised Finance}
One of the major applications of blockchain systems, and Ethereum in particular is decentralized finance, often referred to as DeFi.
DeFi is the development of financial systems on top of blockchains using smart contracts.
DeFi systems have several major characteristics:
\begin{description}
  \setlength{\baselineskip}{15pt}
\item[Non-custodial] Users of DeFi systems should have control over their funds at all time
\item[Permisionless] DeFi systems should be available to everyone
\item[Openly auditable] Anyone has the ability to verify the state of a DeFi system at any point in time
\item[Composable] DeFi systems can be freely composed to interact with one another.
\end{description}

There are several primitive, in the form of smart contracts, that are often used when building DeFi systems.

\subsection{Oracles}
One of the major challenges smart contracts face concerns access to off-chain information, i.e. data that does not natively exist on-chain.
Oracles are data feeds into smart contracts and provide a mechanism for accessing off-chain information through some third party.
In DeFi, oracles are commonly used for price feed data to determine the real-time price of assets.
For instance, via the Compound Open Price Feed~\cite{web:compoundfinance_prices}, vetted third-party reporters sign off on price data using a known public key, where the resulting feed can be relied upon by smart contracts.

\subsection{Stablecoins}
An alternative to volatile cryptoassets is given by stablecoins, which are priced against a peg and can be either custodial or non-custodial.
For custodial stablecoins (e.g. \coin{USDC}~\cite{web:usdc}), tokens represent a claim of some off-chain reserve asset, such as fiat currency, which has been entrusted to a custodian.
Non-custodial stablecoins (e.g. \coin{DAI}~\cite{whitepaper:maker}) seek to establish price stability via economic mechanisms specified by smart contracts.
For a thorough discussion on stablecoin design, we direct the reader to \cite{Klages-Mundt2020}.


\input{5a-smart-contract-security/main}

\input{5b-economic-security/main}
